\documentclass[../../main.tex]{subfiles}
\begin{document}

API for a Regression model object
\begin{enumerate}
    \item \code{construct(parameters)} \\
        Construct/initialise the object with the specified parameters.
    \item \code{fit(X, y)} \\
        Fit the model parameters to the data provided, with inputs X and outputs y.
    \item \code{predict(x)} \\
        Given a new input x, return a prediction of y and the variance of the prediction, given fitted parameters.
\end{enumerate}

API for a Regression manager object
\begin{enumerate}
    \item \code{define(feature\_names, output\_name)} \\
        Receive semantic information about the input features and output, and the number of features expected.
    \item \code{load\_data(X, y)} \\
        Receive a dataset with inputs X and outputs y, to be used for training and testing.
    \item \code{add\_model(model, parameters)} \\
        Add a model to the list of models to use, with the specified parameters.
    \item \code{train()} \\
        Train each model provided by fitting their parameters to the training data. Evaluate the performance of each model against the test data. Determine the best model.
    \item \code{predict(x)} \\
        Given a new input x, return the most likely output, using the model which performed best against the training data.
\end{enumerate}

API for helper functions
\begin{enumerate}
    \item \code{describe\_data(regression\_manager)} \\
        Return a string describing the data loaded into a regression manager.
    \item \code{plot\_data(regression\_manager, args...)} \\
        Plot the data loaded into a regression manager, dependent on some arguments.
    \item \code{describe\_model(regression\_model)} \\
        Return a string describing a given regression model.
    \item \code{summarise\_results(regression\_manager)} \\
        Return a string summarising the results of training the models.
    \item \code{plot\_results(regression\_manager, args...)} \\
        Plot information about the results of training, dependent on some arguments.
\end{enumerate}

\end{document}
